---
title: "IMDB Non-Commercial Data"
author: "Minnie Yuan, Zixuan Wang, Yongyi Yang"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

The [Internet Movie Database (IMDb)]() contains information on millions of movies and television programs. They offer several [non-commercial use datasets](https://developer.imdb.com/non-commercial-datasets/).

These data are quite numerous, included with the this project is a random sample of 100,000 titles (movies/TV programs), along with all all the ratings associated with those titles, all principals (actors) involved in the production, and some basic demographics about those actors. More details on the pre-processing can be found in the `/original` folder in this directory. Due to size constraints, the original data is not included in this repository, but you can down load the `.tsv.gz` files from the link above if you wish to recreate these files.

## Tables

The data will require additional arguments to the `read_csv` to properly handle how missing data is coded. These code chunk is provide as an example only.

```{r cache = TRUE}
title_basics <- read_csv('title_basics_sample.csv.gz')
title_principals <- read_csv('title_principals_sample.csv.gz')
title_ratings <- read_csv('title_ratings_sample.csv.gz')
names_basics <- read_csv('name_basics_sample.csv.gz')
```

## Columns
Please refer to the [IMDb documentation on columns names and their meanings](https://developer.imdb.com/non-commercial-datasets/). Take special note of columns that are used as primary and secondary keys.

```{r}
colnames(title_basics)
colnames(title_principals)
colnames(title_ratings)
colnames(names_basics)
```

#part 2-2
```{r}
table(title_principals$category)
sort(table(names_basics$primaryProfession), decreasing = TRUE)[1:10]
```

#part 2-3

Graph of Names_basics 
```{r}
summary(names_basics)
names_basics %>%
  mutate(knownForCount = lengths(strsplit(knownForTitles, ","))) %>%
  ggplot(aes(knownForCount)) +
  geom_histogram(binwidth = 1, fill = "yellow") +
  labs(title = "Distribution of Known Titles per Person", x = "Number of Titles", y = "Count")
```



Graph of title_principles
```{r}
summary(title_principals)
title_principals %>%
  count(category, sort = TRUE) %>%
  mutate(category = fct_reorder(category, n)) %>%
  ggplot(aes(x = category, y = n)) +
  geom_bar(stat = "identity", fill = "orange") +
  coord_flip() +
  labs(title = "Categories Distributions in Title Principals", x = "Category", y = "Count")
```

#part2-3
#title_basic table group by is adult and titletype
```{r}
summary(title_basics)

# runtimeMinutes by titleType and isAdult
title_basics %>%
  filter(!is.na(runtimeMinutes)) %>%
  ggplot(aes(x = titleType, y = runtimeMinutes, fill = factor(isAdult))) +
  geom_boxplot() +
  coord_flip() +
  labs(title = "Runtime Distribution by Title Type and Adult Content", x = "Title Type", y = "Runtime (minutes)", fill = "Is Adult")

# Count titles with original vs different release names
alt_titles <- title_basics %>%
  filter(originalTitle != primaryTitle)
num_alt_titles <- nrow(alt_titles)
cat("Number of titles known by a different name than original:", num_alt_titles)

#Release year conditional distribution
ggplot(alt_titles, aes(x = startYear, fill = titleType)) +
  geom_density(alpha = 0.5) +
  labs(title = "Release Year Distributions of Titles Known by Different Names", x = "Release Year", fill = "Title Type")
```

7244 titles in your dataset have a primaryTitle that is not the same as the originalTitle.

#part2-4
```{r}
title_basics <- title_basics %>%
  mutate(
    startYear = as.numeric(startYear),  # Convert to numeric
    isAdult = as.factor(isAdult)        # Convert to factor for grouping
  ) %>%
  filter(
    !is.na(startYear),                  # Remove missing years
    !is.na(titleType),                  # Remove missing title types
    !is.na(isAdult)                     # Remove missing adult flags
  )

ggplot(title_basics, aes(x = startYear, fill = titleType)) +
  geom_density(alpha = 0.6, adjust = 1.5) +  # Density plot for smooth distribution
  facet_wrap(~ titleType, ncol = 2) +        # Facet by titleType
  labs(
    title = "Release Year Distribution by Title Type",
    x = "Release Year",
    y = "Density"
  ) +
  theme_minimal() +
  theme(legend.position = "none")            # Remove redundant legend


ggplot(title_basics, aes(x = startYear, fill = isAdult)) +
  geom_histogram(binwidth = 5, position = "dodge", alpha = 0.8) +  # Dodged histogram
  labs(
    title = "Release Year Distribution by Adult Content Flag",
    x = "Release Year",
    y = "Count"
  ) +
  scale_fill_manual(
    name = "Adult Content",
    labels = c("Non-Adult", "Adult"),
    values = c("#1f77b4", "#ff7f0e")          # Custom colors
  ) +
  theme_minimal()
```


#part2-5
```{r}
summary(title_ratings)
# Break average ratings into three groups
title_ratings <- title_ratings %>%
  mutate(ratingGroup = cut(averageRating, breaks = 3, labels = c("Low", "Medium", "High")))

ggplot(title_ratings, aes(x = ratingGroup, y = numVotes)) +
  geom_violin(fill = "purple") +
  labs(title = "Number of Votes by Rating Groups", x = "Rating Group", y = "Number of Votes")
```
Higher rated titles rated more often than lower rated titles

#part 2-6
```{r}
# Count titles each person is known for
names_basics <- names_basics %>%
  mutate(
    num_known_titles = sapply(strsplit(knownForTitles, ","), function(x) sum(x != ""))
  )

# Plot distribution
ggplot(names_basics, aes(x = num_known_titles)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Number of Titles People Are Known For",
       x = "Number of Known Titles",
       y = "Count") +
  theme_minimal() +
  scale_x_continuous(breaks = 0:max(names_basics$num_known_titles))

# Convert years to numeric and calculate age
current_year <- 2025  # From problem context

names_age <- names_basics %>%
  mutate(
    birthYear = as.numeric(birthYear),
    deathYear = as.numeric(deathYear),
    status = ifelse(is.na(deathYear), "Living", "Deceased"),
    age = ifelse(status == "Deceased", 
                deathYear - birthYear, 
                current_year - birthYear)
  ) %>%
  filter(!is.na(birthYear))  # Remove missing birth years

#Decrease cast members age distribution 
ggplot(names_age %>% filter(status == "Deceased"), aes(x = age)) +
  geom_histogram(bins = 20, fill = "salmon", color = "black") +
  labs(title = "Age Distribution of Deceased Cast Members",
       x = "Age at Death",
       y = "Count") +
  theme_minimal()

#Living cast members
ggplot(names_age %>% filter(status == "Living"), aes(x = age)) +
  geom_histogram(bins = 20, fill = "lightgreen", color = "black") +
  labs(title = "Age Distribution of Living Cast Members",
       x = "Current Age",
       y = "Count") +
  theme_minimal()
```

#part 2-7
Find names 
```{r}
# Part 2-7: Find actors with Tom variants (CORRECTED)
toms <- names_basics %>%  # Changed from name_basics to names_basics
  filter(grepl("^(Tom|Thomas|Thom|Tomas)\\b", primaryName, ignore.case = TRUE))

cat("Number of actors named Tom, Thomas, Thom, or Tomas:", nrow(toms))
names_basics <- read_csv('name_basics_sample.csv.gz')  # Plural "names_basics"
```

```{r}
# Function to check alliteration
alliteration_titles <- title_basics %>%
  filter(grepl("^(\\w)\\w*\\s(\\1\\w*\\s?)+$", primaryTitle, ignore.case = TRUE))
cat("Number of titles using alliteration:", nrow(alliteration_titles))
```

# Part 3-1
new version of title_basics
```{r}
library(tidyverse)
title_genres <- title_basics %>%
  separate_rows(genres, sep = ",")
```

#part 3-2 line plot
```{r}
library(tidyverse)
top_genres <- title_genres %>%
  count(genres, sort = TRUE) %>%
  top_n(10, n) %>%
  pull(genres)

title_genres %>%
  filter(genres %in% top_genres) %>%
  group_by(startYear, genres) %>%
  summarize(count = n()) %>%
  ggplot(aes(x = as.numeric(startYear), y = count, color = genres)) +
  geom_line() +
  labs(title = "Count of Top Genres Over Time", x = "Year", y = "Count", color = "Genre")
```

#part 3-3 model.matrix function 
```{r}
genre_matrix <- model.matrix(~ genres - 1, data = title_genres)
correlations <- cor(genre_matrix)

diag(correlations) <- 0
most_common_pair <- which(correlations == max(correlations), arr.ind = TRUE)

most_common_pair_genres <- colnames(correlations)[most_common_pair[1, ]]
cat("Most common genre pair:", most_common_pair_genres[1], "and", most_common_pair_genres[2])
```
